Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools

### 2023-09-26 paper/PaperTRA/PaperTRA.Rmd
Fix cart typo
Minor text tweaks

abstract: | 
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 20% of cart trips could switch to the bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 6,000 tons/day, with benefits over 10 years of €230 million. For the PT leg, the transfer from car avoids of at least 8,500 tons of CO~2~eq emissions per year.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
bibliography: bibliography.bib
# Use csl when using default citation_package for Pandoc citeproc
csl: https://www.zotero.org/styles/springer-lecture-notes-in-computer-science
# citations with arabic numbers is prefered when using natbib
natbiboptions: numbers

@@

According to the latest mobility survey conducted in 2018 [@IMOB], the LMA registered a total of 5.3 million daily trips, with only 0.5% by bicycle. Car modal share was 58.4%, while PT accounted for 15.5%. The number of intra-municipal trips - with origin and destination in the same municipality - amounts to 3.5 million trips, exceeding the number of inter-municipal trips (1.8 million trips) - involving travel between different municipalities. Cars and public transport are the most used modes for intercity trips, with cars being the predominant choice for all journeys.

@@

To achieve the cycling targets set by the Portuguese national cycling strategy for 2025 and 2030 (4% and 10%, respectively) [@ENMAC], the Lisbon's Metropolitan Department of Transport introduced _biclaR_^[see [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt/)], a decision support tool that facilitates the design and development of a metropolitan cycling network [@felix2023].

>>>>

abstract: | 
  A high proportion of car trips can be replaced by a combination of public transit and cycling for the first-and-last mile. This paper estimates the potential for cycling combined with public transit (PT) as a substitute for car trips in the Lisbon metropolitan area and assesses its socio-environmental impacts using open data and open source tools.
  A decision support tool that facilitates the design and development of a metropolitan cycling network was developed (_biclaR_). The social and environmental impacts were assessed using the _HEAT for Cycling_ and the _HEAT as a Service_ tools. The impacts of shifting car trips to PT were also estimated and monetized.
  The results indicate that 20% of car trips could switch to the bicycle + PT combination. Shifting to cycling for the first-and-last mile stages can reduce annual CO~2~eq emissions from 6,000 tons/day, with benefits over 10 years of €230 million. For the PT leg, the transfer from car avoids of at least 8,500 tons of CO~2~eq emissions per year.
  This evidence can support policymakers to prioritize interventions that reduce the reliance on private motor vehicles.
keywords: 
  - Active transport
  - Intermodality
  - First and last mile
  - Health economic assessment
  - Environmental impacts
  - Open data and methods
bibliography: bibliography.bib
# Use csl when using default citation_package for Pandoc citeproc
csl: https://www.zotero.org/styles/springer-lecture-notes-in-computer-science
# citations with arabic numbers is prefered when using natbib
natbiboptions: numbers

@@

According to the latest mobility survey conducted in 2018 [@IMOB], the LMA registered a total of 5.3 million daily trips, with only 0.5% by bicycle. Car modal share was 58.4%, while PT accounted for 15.5%. The number of intra-municipal trips --- with origin and destination in the same municipality --- amounts to 3.5 million trips. This exceeds the number of inter-municipal trips (1.8 million trips), involving travel between different municipalities. Cars and public transport are the most used modes for intercity trips, with cars being the predominant choice for all journeys.

@@

To achieve the cycling targets set by the Portuguese national cycling strategy for 2025 and 2030 (4% and 10%, respectively) [@ENMAC], the Lisbon's Metropolitan Department of Transport introduced _biclaR_^[See [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt/)], a decision support tool that facilitates the design and development of a metropolitan cycling network [@felix2023].
